---
title: "R Notebook"
output: html_notebook
---

I'm going to write this one blind

## Libraries
```{r}
library(tidyverse)
library(lubridate)
library(cowplot)
library(survminer)
```

# Import the data

- import the data and list the contents

```{r}
tuesdata <- tidytuesdayR::tt_load('2020-04-28')

tuesdata %>% names()

```

THere are 4 tables

lets explore them

Print off the head of each table

```{r}
tuesdata %>% map(head)

tuesdata %>% map(dim)

grosses <- tuesdata$grosses
synopses <- tuesdata$synopses
cpi <- tuesdata$cpi
pre1985 <- tuesdata$`pre-1985-starts`


```

Ah, so this looks to be about Broadway theatre!.
- Grosses shows weekly information on various shows including their gross, ticket prices and theatre capacity.
- Synopsis contains text summarising each show
- CPI contains inflation data so we can convert historic values to todays value
- pre-1985 starts contains start data - presumably this is for shows that started before we havce the other data

## Ideas

- what were the top shows of each year
- Has seat capacity gone up
- How has highest and average ticket price changes - looks like top ticet price changes 
- Capacity vs average ticket price vs gross sweet spots
- Forcast correction based on annual fluctuations - more tickets sold over holidays

# Top grossing shows per month

- we need to calc ave cpi per year

```{r}
cpi_per_year <- cpi %>% mutate(year = year(year_month)) %>%
  group_by(year) %>%
  summarise(cpi = max(cpi))

cpi <- cpi %>% mutate(year = year(year_month), month = month(year_month)) %>% mutate(multiplier =max(cpi)/cpi)


grosses %>%
  mutate(year = year(week_ending)) %>%
  filter(year != 2020) %>%
  group_by(show,year) %>%
    summarise(yearly_gross = sum(weekly_gross)) %>%
  left_join(cpi_per_year) %>%
  mutate(yearly_gross_today = yearly_gross *cpi) %>%
  ungroup() %>%
  ggplot(aes(year, yearly_gross_today/100, group = year)) + 
  geom_boxplot() +
  theme_cowplot() +
  labs(title = "CPI adjusted annual gross for the top perfoming shows has increaced",
       x= "Year", y= "CPI adjusted annual gross")


```

# Can we see a weekly change accross years

```{r}
grosses_adjusted <- grosses %>%
  mutate(year = year(week_ending),
         month = month(week_ending)) %>%
  left_join(cpi) %>%
  mutate_at(vars(weekly_gross_overall, weekly_gross:top_ticket_price), ~.x*multiplier)

grosses_adjusted %>% ggplot(aes(week_number, weekly_gross, group = show)) + 
  geom_line(alpha = 0.1) + 
  facet_wrap(~year, scales = "free_y") +
  theme_cowplot() +
  theme(axis.text.y = element_blank())

grosses_adjusted %>% filter(year == 2020)
```

Week number corresponds to broadway seasons not year, so we need to change year to season

```{r}
grosses_adjusted %>% 
  pivot_longer(cols = weekly_gross:previews, names_to = "metric", values_to = "value") %>%
  group_by(year_month, metric) %>%
  summarise(mean = median(value, na.rm = TRUE)) %>%
  ggplot(aes(year_month, mean)) + geom_point() +
  facet_wrap(~metric, scales = "free_y")

grosses_adjusted %>%
  group_by(show,year) %>%
    summarise(ave_ticket_price = mean(avg_ticket_price)) %>%
  ggplot(aes(year, ave_ticket_price, group = year)) + 
  geom_boxplot() +
  theme_cowplot() +
  labs(title = "CPI adjusted average ticket price for the top perfoming shows has increaced",
       x= "Year", y= "CPI adjusted average ticket price")


grosses_adjusted %>%
  group_by(show,year) %>%
    summarise(ave_ticket_price = mean(avg_ticket_price/100)) %>% arrange(desc(ave_ticket_price))

grosses_adjusted %>% filter(year == 2019) %>% ggplot(aes(avg_ticket_price, pct_capacity)) + geom_point() +
  scale_x_log10()

grosses_adjusted %>% filter(pct_capacity <= 0.7& year == 2019) %>%
  group_by(show) %>%
  summarise(avg_ticket_price = mean(avg_ticket_price), pct_capacity = mean(pct_capacity)) %>%
  arrange(desc(avg_ticket_price))

cpi %>% ggplot(aes(year_month, cpi)) + geom_point()

```

# show survival analysis

```{r warning=FALSE}
library(survival)

grosses_adjusted %>% arrange(desc(week_ending))

show_run_length <- grosses_adjusted %>%
  filter(!show %in% pre1985$show) %>%
  select(show, theatre, week_ending,previews, performances, pct_capacity) %>%
  arrange(show,theatre,week_ending) %>%
  mutate(group = ifelse(week_ending - lag(week_ending) >= weeks(8) | show != lag(show) | theatre != lag(theatre), 1,0),
         group = replace_na(group,1),
         group = cumsum(group) + 1) %>%
  group_by(show, theatre, group) %>%
  summarise(opening_week = min(week_ending),
         end_week = max(week_ending),
         performances = sum(performances) + sum(previews),
         mean_fill = mean(pct_capacity)
         ) %>%
  mutate(closed = end_week != date("2020-03-01")) %>%
  mutate(length_of_run =  interval(opening_week, end_week) %/% days(1))

show_run_length %>% arrange(desc(length_of_run)) %>% mutate(years = length_of_run/365) %>%
  ggplot(aes(performances, length_of_run)) + geom_point(shape = ".") + scale_x_log10() + scale_y_log10()

show_run_length %>% arrange(desc(length_of_run)) %>% 
  mutate(years = length_of_run/12, weeks = years*52, spw = performances/weeks) %>% 
  ggplot(aes(spw)) + geom_histogram()

show_run_length %>% arrange(desc(length_of_run)) %>% 
  mutate(years = length_of_run/12, weeks = years*52, spw = performances/weeks) %>% 
  filter(spw <= 5)

show_survival <- show_run_length %>%
  survfit(Surv(performances, closed) ~1, data = .) 

show_survival
plot(show_survival)

survival_tidy <- tibble(performances = show_survival$time, 
                        prob = show_survival$surv)

survival_tidy %>% ggplot(aes(time, prob)) + geom_line() + scale_x_log10()

show_survival %>% names()

show_run_length %>% filter(performances <= 112)

show_run_length %>% ungroup() %>%distinct(show, group, performances) %>% count(show, sort = TRUE)

show_run_length %>% ggplot(aes(mean_fill)) + geom_histogram()

```

Do shold out shows have longer runs

```{r}
show_run_length %>% mutate(closed = ifelse(performances >= 1000, FALSE, TRUE),
                           performances = ifelse(performances >= 1000, 1000, performances)) %>%
  mutate(sellout = mean_fill >=0.75) %>% 
survfit(Surv(performances, closed) ~ sellout, data = .) %>% 
  plot()

plot(show_survival)



survival_tidy <- tibble(performances = show_survival$time, 
                        prob = show_survival$surv)

survival_tidy %>% ggplot(aes(time, prob)) + geom_line() + scale_x_log10()
```


How to calculate opening week fill

```{r}


grosses_adjusted %>%
  filter(!show %in% pre1985$show) %>%
  select(show, theatre, week_ending,previews, performances, pct_capacity) %>%
  arrange(show,theatre,week_ending) %>%
  mutate(group = ifelse(week_ending - lag(week_ending) >= weeks(8) | show != lag(show) | theatre != lag(theatre), 1,0),
         group = replace_na(group,1),
         group = cumsum(group) + 1) %>%
  group_by(show, theatre, group) %>%
  transmute(opening_week = min(week_ending),
         end_week = max(week_ending),
         performances = sum(performances) + sum(previews),
         opening_week_fill = ifelse(week_ending == opening_week, pct_capacity, NA)) %>%
  na.omit() %>%
  mutate(closed = ifelse(performances >= 1000, FALSE, TRUE),
                           performances = ifelse(performances >= 1000, 1000, performances)) %>%
  mutate(sellout = opening_week_fill >=1) %>% 
survfit(Surv(performances, closed) ~ sellout, data = .) %>% 
  plot()

grosses_adjusted %>%
  filter(!show %in% pre1985$show) %>%
  arrange(show,theatre,week_ending) %>%
  mutate(group = ifelse(week_ending - lag(week_ending) >= weeks(8) | show != lag(show) | theatre != lag(theatre), 1,0),
         group = replace_na(group,1),
         group = cumsum(group) + 1) %>%
  group_by(show, theatre, group) %>%
  transmute(opening_week = min(week_ending),
         end_week = max(week_ending),
         performances = sum(performances) + sum(previews),
         opening_week_fill = ifelse(week_ending == opening_week, pct_capacity, NA),
         opening_week_gross = ifelse(week_ending == opening_week, weekly_gross, NA),
         seats_in_theatre = seats_in_theatre) %>%
  na.omit() %>%
  mutate(closed = ifelse(performances >= 1000, FALSE, TRUE),
                           performances = ifelse(performances >= 1000, 1000, performances)) %>%
  ungroup() %>%
  mutate(seats_bin = ntile(seats_in_theatre, 5)) %>% 
survfit(Surv(performances, closed) ~ seats_bin, data = .) %>% 
  plot()
```

```{r}
show_run_data <- grosses_adjusted %>%
  filter(!show %in% pre1985$show) %>%
  arrange(show,theatre,week_ending) %>%
  mutate(group = ifelse(week_ending - lag(week_ending) >= weeks(8) | show != lag(show) | theatre != lag(theatre), 1,0),
         group = replace_na(group,1),
         group = cumsum(group) + 1) %>%
  group_by(show, theatre, group) %>%
  transmute(opening_week = min(week_ending),
         end_week = max(week_ending),
         performances = sum(performances) + sum(previews),
         week1_fill = ifelse(week_ending == opening_week, pct_capacity, NA),
         week1_gross = ifelse(week_ending == opening_week, weekly_gross, NA),
         week1_ave_ticket_price = ifelse(week_ending == opening_week, avg_ticket_price, NA),
         seats_in_theatre = seats_in_theatre) %>%
  na.omit() %>%
  mutate(closed = ifelse(performances >= 500 | end_week == date("2020-03-01"), FALSE, TRUE),
         performances = ifelse(performances >= 500, 500, performances)) %>%
  group_by(show) %>%
  arrange(opening_week) %>%
  mutate(revival = opening_week > min(opening_week),
         revival_num = c(1:n())) %>%
  ungroup()

```

# week1 ticket price

```{r, fig.height=6, fig.width=4}
subset <- show_run_data %>% 
  mutate(bin = ntile(week1_ave_ticket_price, 3),
         group = case_when(bin == 1 ~ "low",
                           bin == 2 ~ "mid",
                           bin == 3 ~ "high"))

fit <- survfit(Surv(performances, closed) ~ group, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           fun = "pct",
           risk.table = TRUE)
```
# week 1 capacity
```{r, fig.height=6, fig.width=4}
subset <- show_run_data %>% 
  mutate(bin = ntile(week1_fill, 3),
         group = case_when(bin == 1 ~ "low",
                           bin == 2 ~ "mid",
                           bin == 3 ~ "high"))

fit <- survfit(Surv(performances, closed) ~ group, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           fun = "pct",
           risk.table = TRUE)
```

# week1 gross

```{r, fig.height=6, fig.width=6}
subset <- show_run_data %>% 
  mutate(bin = ntile(week1_gross, 3),
          group = case_when(bin == 1 ~ "low",
                           bin == 2 ~ "mid",
                           bin == 3 ~ "high"),
         group = as.factor(group),
         group = fct_reorder(group, bin)
         )

fit <- survfit(Surv(performances, closed) ~ group, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           fun = "pct",
           risk.table = TRUE,
           legend.title = "Opening week gross",
           legend.labs = c("Low", "Medium", "High"),
           surv.median.line = "hv",
           tables.height = 0.3)
```
# theatre_size

```{r, fig.width=6, fig.height=6}
subset <- show_run_data %>% 
  mutate(bin = ntile(seats_in_theatre, 2),
         group = case_when(bin == 1 ~ "low",
                           #bin == 2 ~ "mid",
                           bin == 2 ~ "high"),
         group = as.factor(group),
         group = fct_reorder(group, bin)
         )

fit <- survfit(Surv(performances, closed) ~ group + revival, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           fun = "pct",
           risk.table = TRUE,
           legend.title = "Size of theatre",
           #legend.labs = c("Small", "Medium", "Large"),
           surv.median.line = "hv",
           tables.height = 0.3)

aareg(Surv(performances, closed) ~ group + revival, data = subset) %>% plot()
```

# revivals

```{r, fig.height=6, fig.width=6}


fit <- survfit(Surv(performances, closed) ~ revival, data = show_run_data) 


ggsurvplot(fit, show_run_data,
           conf.int = T,
           pval = TRUE,
           fun = "pct",
           risk.table = TRUE)

autoplot(fit)

```

# Can we pull something from text analysis

```{r}

synopses %>% 
  filter(grepl("film|movie",synopsis)) %>%
  sample_n(5) %>%
  pull(synopsis)

synopses %>% 
  filter(grepl("Tony",synopsis)) %>%
  sample_n(5) %>%
  pull(synopsis)



```
### Do musicals mentioning film or movies have longer runs?

```{r}
subset <- show_run_data %>% 
  left_join(synopses) %>%
  mutate(film = grepl("film|movie",synopsis))

fit <- survfit(Surv(performances, closed) ~ film, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           fun = "pct",
           risk.table = TRUE,
           legend.title = "Synopsis mentions film or movie",
           #legend.labs = c("Small", "Medium", "Large"),
           surv.median.line = "hv",
           tables.height = 0.3)
```

Do movies mentioning Tony survivie longr

```{r}
subset <- show_run_data %>% 
  left_join(synopses) %>%
  mutate(tony = grepl("Tony",synopsis))

fit <- survfit(Surv(performances, closed) ~ tony, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           fun = "pct",
           risk.table = TRUE,
           legend.title = "Synopsis mentions Tony",
           #legend.labs = c("Small", "Medium", "Large"),
           surv.median.line = "hv",
           tables.height = 0.3)
```



Sentiment analysis?

```{r}
library(tidytext)

data("stop_words")
sentiments <- get_sentiments("afinn")


show_sentiment <- synopses %>%
  unnest_tokens(word, synopsis) %>%
  anti_join(stop_words) %>%
  inner_join(sentiments) %>% group_by(show) %>%
  summarise(n = n(),
            median = median(value, na.rm = TRUE),
            sd = sd(value, na.rm = TRUE)) %>%
  arrange(desc(median)) %>%
  mutate(sentiment = case_when(
    median >=2 & n > 3 &median-sd > 0~ "happy",
    median <= -2 & n > 3 & median+sd <0 ~ "sad",
    TRUE ~ "unknown"
  ))
  

show_sentiment %>% top_n(1, median) %>%
  left_join(synopses) %>%
  pull(synopsis)

show_sentiment %>% top_n(-1, median) %>%
  left_join(synopses) %>%
  pull(synopsis)

show_sentiment %>% ggplot(aes(median)) + geom_histogram()

show_sentiment %>% #filter(sentiment != "unknown") %>%
  mutate(show = factor(show),
         show = fct_reorder(show, median)) %>%
  ggplot(aes(median, show)) +
  geom_pointrange(aes(xmin = median-sd, xmax = median+sd, colour = sentiment)) +
  theme(axis.text.y = element_blank())+ theme_void() +
  geom_vline(xintercept = 0)

show_sentiment

```

Plot surfvial analysis based on this

```{r}
subset <- show_run_data %>% 
  left_join(show_sentiment) %>%
  filter(sentiment != "unknown")

fit <- survfit(Surv(performances, closed) ~ sentiment, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           risk.table = TRUE,
           legend.title = "Synopsis mentions film or movie",
           #legend.labs = c("Small", "Medium", "Large"),
           surv.median.line = "hv",
           tables.height = 0.3)
```

What about a differnt sentiment library

```{r}
nrc_sentimnets <- get_sentiments("nrc")

show_sentiment_nrc <- synopses %>%
  unnest_tokens(word, synopsis) %>%
  anti_join(stop_words) %>%
  inner_join(nrc_sentimnets) %>% 
  group_by(show) %>%
  summarise(n = n(),
            n_pos = sum(sentiment == "positive"),
            n_neg = sum(sentiment == "negative"),
            score = (n_pos-n_neg)/(n_pos+n_neg))

show_sentiment_nrc <-synopses %>%
  unnest_tokens(word, synopsis) %>%
  anti_join(stop_words) %>%
  inner_join(nrc_sentimnets) %>% 
  group_by(show) %>%
  count(sentiment) %>%
  filter(!sentiment %in% c("positive", "negative")) %>%
  top_n(1,n)

show_sentiment_nrc %>% arrange(desc(score))
```
```{r, fig.height=8, fig.width=8}
`
subset <- show_run_data %>% 
  left_join(show_sentiment_nrc)

fit <- survfit(Surv(performances, closed) ~ sentiment, data = subset) 


ggsurvplot(fit, subset,
           conf.int = F,
           pval = TRUE,
           risk.table = TRUE,
           legend.title = "Synopsis modal sentiment",
           #legend.labs = c("Small", "Medium", "Large"),
           surv.median.line = "hv",
           tables.height = 0.3)
```


```{r}
subset <- show_run_data %>% 
  left_join(show_sentiment_nrc) %>%
  filter(sentiment %in% c("disgust", "joy", "anticipation"))

fit <- survfit(Surv(performances, closed) ~ sentiment, data = subset) 


ggsurvplot(fit, subset,
           conf.int = T,
           pval = TRUE,
           risk.table = TRUE,
           legend.title = "Synopsis modal sentiment",
           legend.labs = c("Anticipation", "Disgust", "Joy"),
           surv.median.line = "hv",
           tables.height = 0.3)
```
```

