---
title: "Week 48 - Student loan payments"
output: html_notebook
---
https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-11-26

# Dependencies
```{r}
library(tidyverse)
library(lubridate)
library(cowplot)
```


# Get the data

```{r}
loans <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-11-26/loans.csv")

```

## Initial inspection

```{r}
loans

loans %>% filter(agency_name == "Van Ru Credit Corporation") 
```

Loans contains a tibble of loans with 291 observations and 10 variables.
- Agency name is a chr
- Year is a 2 digit dbl
- Quater is a 2 digit dbl
- Others are all double
- Some information from added is missing. This is the total value added during the quarter. If no value was addied is this na?


# Initial vis

## Plot the total debt over time

```{r}
loans %>% 
  mutate(yq  =paste0("20", year, "\nQ", quarter)) %>%
  filter(year != 15) %>%
  select(-year, -quarter) %>%
  group_by(yq) %>%
  pivot_longer(starting:total, names_to = "metric", values_to = "value") %>%
  mutate(value = value/1e6) %>%
  arrange(value) %>%
  ggplot(aes(yq, value)) +
  geom_point(aes(group = value), position=position_dodge(width=0.75), size = 0.5, alpha = 0.5) +
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "crossbar", width = 0.5, size = 0.25) +
  stat_summary(fun.y = quantile, fun.ymin = quantile, fun.ymax = quantile,
                 geom = "crossbar", width = 0.25, size = 0.25, colour = "red",
               fun.args = list(probs = c(0.25,0.75))) +
  
  #ylim(0,1.2e5) +
  theme(legend.position = "none") +
  theme_cowplot() +
  ylab("Value\nUSD (millions)") +
  facet_wrap(~metric, scales = "free_y") +
  scale_x_discrete(breaks = c("2016\nQ1", "2018\nQ1", "2017\nQ1"))

  
```

Plot the distribution of loan values
```{r}

loans %>% 
  mutate(yq  =paste0("20", year, "\nQ", quarter)) %>%
  filter(year != 15) %>%
  pivot_longer(starting:total, names_to = "metric", values_to = "value") %>%
  mutate(value = value/1e6) %>%
  filter(year != 15) %>%
  select(-year, -quarter) %>%
  ggplot(aes(value)) +
  geom_density() +
  facet_grid(yq~metric, scales = "free")


```






# Look by quater grouo by year

```{r}
loans %>% 
  mutate(yq  =paste(year, quarter, sep = ".")) %>%
  ggplot(aes(quarter, total, colour = agency_name, group = agency_name)) + 
  geom_line() + 
  geom_point() +
  theme(legend.position = "none") +
  facet_wrap(~year)
```


# UMAP

```{r}
loans %>% select(-agency_name, year, quarter) %>% mutate_all(replace_na, 0) %>%
  uwot::umap() %>%
  as_tibble() %>%
  rename(umap1 = V1, umap2 = V2) %>%
  bind_cols(loans) %>%
  ggplot(aes(umap1, umap2, colour = total)) +
  geom_point(size = 0.05) +
  coord_equal() +
  theme_cowplot()
```

